Northern Pikeminnow Abundance in Deadwater Slough, Salmon River, Idaho and Potential Impacts to Local Chinook Salmon Populations

### 2022-02-01 analysis/paper/paper.Rmd
added bioenergetics table

# fall, fish 10% - 90%
fall_f10 = read_csv(here("analysis/data/raw_data/Bioenergetics_Fall_I90_F10.csv")) %>% fmt_bio_out()
fall_f20 = read_csv(here("analysis/data/raw_data/Bioenergetics_Fall_I80_F20.csv")) %>% fmt_bio_out()
fall_f30 = read_csv(here("analysis/data/raw_data/Bioenergetics_Fall_I70_F30.csv")) %>% fmt_bio_out()
fall_f40 = read_csv(here("analysis/data/raw_data/Bioenergetics_Fall_I60_F40.csv")) %>% fmt_bio_out()
fall_f50 = read_csv(here("analysis/data/raw_data/Bioenergetics_Fall_I50_F50.csv")) %>% fmt_bio_out()
fall_f60 = read_csv(here("analysis/data/raw_data/Bioenergetics_Fall_I40_F60.csv")) %>% fmt_bio_out()
fall_f70 = read_csv(here("analysis/data/raw_data/Bioenergetics_Fall_I30_F70.csv")) %>% fmt_bio_out()
fall_f80 = read_csv(here("analysis/data/raw_data/Bioenergetics_Fall_I20_F80.csv")) %>% fmt_bio_out()
fall_f90 = read_csv(here("analysis/data/raw_data/Bioenergetics_Fall_I10_F90.csv")) %>% fmt_bio_out()

# spring, fish 10% - 90%
spring_f10 = read_csv(here("analysis/data/raw_data/Bioenergetics_Spring_I90_F10.csv")) %>% fmt_bio_out()
spring_f20 = read_csv(here("analysis/data/raw_data/Bioenergetics_Spring_I80_F20.csv")) %>% fmt_bio_out()
spring_f30 = read_csv(here("analysis/data/raw_data/Bioenergetics_Spring_I70_F30.csv")) %>% fmt_bio_out()
spring_f40 = read_csv(here("analysis/data/raw_data/Bioenergetics_Spring_I60_F40.csv")) %>% fmt_bio_out()
spring_f50 = read_csv(here("analysis/data/raw_data/Bioenergetics_Spring_I50_F50.csv")) %>% fmt_bio_out()
spring_f60 = read_csv(here("analysis/data/raw_data/Bioenergetics_Spring_I40_F60.csv")) %>% fmt_bio_out()
spring_f70 = read_csv(here("analysis/data/raw_data/Bioenergetics_Spring_I30_F70.csv")) %>% fmt_bio_out()
spring_f80 = read_csv(here("analysis/data/raw_data/Bioenergetics_Spring_I20_F80.csv")) %>% fmt_bio_out()
spring_f90 = read_csv(here("analysis/data/raw_data/Bioenergetics_Spring_I10_F90.csv")) %>% fmt_bio_out()

# compile fall results
fall_results = fall_f10 %>%
  select(date, temperature_c, cum_cons_fish_g) %>%
  rename(f10_g = cum_cons_fish_g) %>%
  mutate(f20_g = fall_f20$cum_cons_fish_g,
         f30_g = fall_f30$cum_cons_fish_g,
         f40_g = fall_f40$cum_cons_fish_g,
         f50_g = fall_f50$cum_cons_fish_g,
         f60_g = fall_f60$cum_cons_fish_g,
         f70_g = fall_f70$cum_cons_fish_g,
         f80_g = fall_f80$cum_cons_fish_g,
         f90_g = fall_f90$cum_cons_fish_g) %>%
  mutate(season = "Fall") #%>%
  # mutate( across(contains('_g'),
  #                .fns = list(n = ~./avg_DSR_g),
  #                .names = "{col}_{fn}") )

# compile spring results
spring_results = spring_f10 %>%
  select(date, temperature_c, cum_cons_fish_g) %>%
  rename(f10_g = cum_cons_fish_g) %>%
  mutate(f20_g = spring_f20$cum_cons_fish_g,
         f30_g = spring_f30$cum_cons_fish_g,
         f40_g = spring_f40$cum_cons_fish_g,
         f50_g = spring_f50$cum_cons_fish_g,
         f60_g = spring_f60$cum_cons_fish_g,
         f70_g = spring_f70$cum_cons_fish_g,
         f80_g = spring_f80$cum_cons_fish_g,
         f90_g = spring_f90$cum_cons_fish_g) %>%
  mutate(season = "Spring") # %>%
  # mutate( across(contains('_g'),
  #                .fns = list(n = ~./avg_NRR_g),
  #                .names = "{col}_{fn}") )

bio_results = rbind(fall_results, spring_results)
bio_g_summ = bio_results %>%
  select(season, f10_g:f90_g) %>%
  pivot_longer(f10_g:f90_g) %>%
  group_by(season, name) %>%
  summarize(cumm_g = max(value))

rm(fall_results,
   fall_f10, 
   fall_f20, 
   fall_f30, 
   fall_f40, 
   fall_f50, 
   fall_f60, 
   fall_f70, 
   fall_f80, 
   fall_f90,
   spring_results,
   spring_f10, 
   spring_f20, 
   spring_f30, 
   spring_f40, 
   spring_f50, 
   spring_f60, 
   spring_f70, 
   spring_f80, 
   spring_f90)

@@

The proportion of fish in the Northern Pikeminnow's diet relative to non-fish prey items is unknown in Deadwater Slough. Therefore, we conducted separate model runs that assume different proportions of fish in the diet ranging from 30 - 90%, in 10% increments. These values were supported by observations elsewhere in the Columbia River basin (54-86%: @Shively1996; 48-86%: @Zimmerman1999; 37%: @Gray2001) that suggest fish are the majority of Northern Pikeminnow prey. Fourteen model runs were completed (7 diet scenarios and 2 seasons), each providing an estimate of the total grams of fish consumed by an individual Northern Pikeminnow (***table reference***). To estimate the total biomass of fish consumed by Northern Pikeminnow in Deadwater Slough, we multiplied the grams of fish consumed by an individual Northern Pikeminnow by the estimated Northern Pikeminnow population sizes during fall and spring.

@@

@@

pFish = bio_results %>%
  ggplot(aes(ymd, f50_g)) +
  geom_line(size = 1.1) +
  geom_ribbon(data = bio_results,
              aes(ymd,ymin = f10_g,
                  ymax = f90_g),
              color = "grey55",
              alpha = 0.4) +  theme_bw() +
  labs(y = "Fish consumed (g)",
       x = "", title = "") +
  facet_rep_wrap(~season,scales = "free_x",
                 repeat.tick.labels = F) +
  theme(strip.background = element_rect(color = "white",
                                        fill = "white"),
        strip.text.x = element_text(size = 12),
        axis.text.x = element_blank(),
        axis.text.y = element_text(size = 11,
                                   color = "black"),
        axis.title.y = element_text(vjust = 1,
                                    size = 12),
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.margin = unit(c(-.1, .3, -.3, .2), "cm"))

>>>>

# fall, fish 30% - 90%
fall_f30 = read_csv(here("analysis/data/raw_data/Bioenergetics_Fall_I70_F30.csv")) %>% fmt_bio_out()
fall_f40 = read_csv(here("analysis/data/raw_data/Bioenergetics_Fall_I60_F40.csv")) %>% fmt_bio_out()
fall_f50 = read_csv(here("analysis/data/raw_data/Bioenergetics_Fall_I50_F50.csv")) %>% fmt_bio_out()
fall_f60 = read_csv(here("analysis/data/raw_data/Bioenergetics_Fall_I40_F60.csv")) %>% fmt_bio_out()
fall_f70 = read_csv(here("analysis/data/raw_data/Bioenergetics_Fall_I30_F70.csv")) %>% fmt_bio_out()
fall_f80 = read_csv(here("analysis/data/raw_data/Bioenergetics_Fall_I20_F80.csv")) %>% fmt_bio_out()
fall_f90 = read_csv(here("analysis/data/raw_data/Bioenergetics_Fall_I10_F90.csv")) %>% fmt_bio_out()

# spring, fish 30% - 90%
spring_f30 = read_csv(here("analysis/data/raw_data/Bioenergetics_Spring_I70_F30.csv")) %>% fmt_bio_out()
spring_f40 = read_csv(here("analysis/data/raw_data/Bioenergetics_Spring_I60_F40.csv")) %>% fmt_bio_out()
spring_f50 = read_csv(here("analysis/data/raw_data/Bioenergetics_Spring_I50_F50.csv")) %>% fmt_bio_out()
spring_f60 = read_csv(here("analysis/data/raw_data/Bioenergetics_Spring_I40_F60.csv")) %>% fmt_bio_out()
spring_f70 = read_csv(here("analysis/data/raw_data/Bioenergetics_Spring_I30_F70.csv")) %>% fmt_bio_out()
spring_f80 = read_csv(here("analysis/data/raw_data/Bioenergetics_Spring_I20_F80.csv")) %>% fmt_bio_out()
spring_f90 = read_csv(here("analysis/data/raw_data/Bioenergetics_Spring_I10_F90.csv")) %>% fmt_bio_out()

# compile fall results
fall_results = fall_f30 %>%
  select(date, temperature_c, cum_cons_fish_g) %>%
  rename(f30_g = cum_cons_fish_g) %>%
  mutate(f40_g = fall_f40$cum_cons_fish_g,
         f50_g = fall_f50$cum_cons_fish_g,
         f60_g = fall_f60$cum_cons_fish_g,
         f70_g = fall_f70$cum_cons_fish_g,
         f80_g = fall_f80$cum_cons_fish_g,
         f90_g = fall_f90$cum_cons_fish_g) %>%
  mutate(season = "Fall") #%>%
  # mutate( across(contains('_g'),
  #                .fns = list(n = ~./avg_DSR_g),
  #                .names = "{col}_{fn}") )

# compile spring results
spring_results = spring_f30 %>%
  select(date, temperature_c, cum_cons_fish_g) %>%
  rename(f30_g = cum_cons_fish_g) %>%
  mutate(f40_g = spring_f40$cum_cons_fish_g,
         f50_g = spring_f50$cum_cons_fish_g,
         f60_g = spring_f60$cum_cons_fish_g,
         f70_g = spring_f70$cum_cons_fish_g,
         f80_g = spring_f80$cum_cons_fish_g,
         f90_g = spring_f90$cum_cons_fish_g) %>%
  mutate(season = "Spring") # %>%
  # mutate( across(contains('_g'),
  #                .fns = list(n = ~./avg_NRR_g),
  #                .names = "{col}_{fn}") )

bio_results = rbind(fall_results, spring_results)
bio_g_summ = bio_results %>%
  select(season, f30_g:f90_g) %>%
  pivot_longer(f30_g:f90_g) %>%
  group_by(season, name) %>%
  summarize(cumm_g = max(value))

rm(fall_results,
   fall_f30, 
   fall_f40, 
   fall_f50, 
   fall_f60, 
   fall_f70, 
   fall_f80, 
   fall_f90,
   spring_results,
   spring_f30, 
   spring_f40, 
   spring_f50, 
   spring_f60, 
   spring_f70, 
   spring_f80, 
   spring_f90)

@@

The proportion of fish in the Northern Pikeminnow's diet relative to non-fish prey items is unknown in Deadwater Slough. Therefore, we conducted separate model runs that assume different proportions of fish in the diet ranging from 30 - 90%, in 10% increments. These values were supported by observations elsewhere in the Columbia River basin (54-86%: @Shively1996; 48-86%: @Zimmerman1999; 37%: @Gray2001) that suggest fish are the majority of Northern Pikeminnow prey. Fourteen model runs were completed (7 diet scenarios and 2 seasons), each providing an estimate of the total grams of fish consumed by an individual Northern Pikeminnow (Table \@ref(tab:bio-tbl)). To estimate the total biomass of fish consumed by Northern Pikeminnow in Deadwater Slough, we multiplied the grams of fish consumed by an individual Northern Pikeminnow by the estimated Northern Pikeminnow population sizes during fall and spring.

@@

```{r bio-tbl}
bio_tbl = bio_g_summ %>%
  mutate(`Date Range` = season) %>%
  select(season,
         `Date Range`,
         name,
         cumm_g) %>%
  mutate(name = str_sub(name, 2, 3)) %>%
  rename(Season = season,
         `% Fish in Diet` = name,
         `g Consumed` = cumm_g) %>%
  mutate(`Date Range` = recode(`Date Range`,
                               Fall = "Sep 15 - Nov 30",
                               Spring = "Mar 1 - May 31")) %>%
  kable(booktabs = T,
        caption = "Bioenergetics model runs to estimate the grams of fish consumed during the DSR fall and NRR spring Chinook Salmon emigrations. The date ranges for each model run, the assumed proportion of fish in the diet, and the total grams of fish consumed by an individual Northern Pikeminnow to maintain its body weight for a given season are shown. The Northern Pikeminnow start/end weights were 598.7 and 430.7 g for the fall and spring seasons, respectively.") %>%
  kable_styling()

bio_tbl

@@

pFish = bio_results %>%
  ggplot(aes(ymd, f60_g)) +
  geom_line(size = 1.1) +
  geom_ribbon(data = bio_results,
              aes(ymd,ymin = f30_g,
                  ymax = f90_g),
              color = "grey55",
              alpha = 0.4) +  theme_bw() +
  labs(y = "Fish consumed (g)",
       x = "", title = "") +
  facet_rep_wrap(~season,scales = "free_x",
                 repeat.tick.labels = F) +
  theme(strip.background = element_rect(color = "white",
                                        fill = "white"),
        strip.text.x = element_text(size = 12),
        axis.text.x = element_blank(),
        axis.text.y = element_text(size = 11,
                                   color = "black"),
        axis.title.y = element_text(vjust = 1,
                                    size = 12),
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.margin = unit(c(-.1, .3, -.3, .2), "cm"))
